Reliable error estimation for Sobol' indices

### 2016-08-18 article/template.tex
added template journal and aux. files

\documentclass[]{elsarticle}
\setlength{\marginparwidth}{0.5in}
\usepackage{amsmath,amssymb,amsthm,natbib,
mathtools,bbm,extraipa,mathabx,graphicx,algorithm}
\usepackage{algpseudocode}
%accents,

\newtheorem{lem}{Lemma}
\newtheorem{remark}{Remark}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{algo}{Algorithm}

\newcommand{\fudge}{\fC}
\newcommand{\cc}{\mathcal{C}}
\newcommand{\dtf}{\textit{\doubletilde{f}}}
\newcommand{\cube}{[0,1)^d}
%\renewcommand{\bbK}{\natzero^d}
\newcommand{\rf}{\mathring{f}}
\newcommand{\rnu}{\mathring{\nu}}
\newcommand{\natm}{\naturals_{0,m}}
\newcommand{\wcS}{\widecheck{S}}
\newcommand{\tol}{\text{tol}}
\newcommand{\e}{\text{e}}
\newcommand{\bvec}[1]{\boldsymbol{#1}}
\newcommand{\vx}{\bvec{x}}
\newcommand{\vI}{\bvec{I}}
\newcommand{\vk}{\bvec{k}}
\newcommand{\vw}{\bvec{w}}
\newcommand{\vz}{\bvec{z}}
\newcommand{\vt}{\bvec{t}}
\newcommand{\dif}{\mathsf{d}}
\newcommand{\hf}{\hat{f}}
\newcommand{\hS}{\widehat{S}}
\newcommand{\tS}{\widetilde{S}}
\newcommand{\tf}{\tilde{f}}
\newcommand{\fC}{\mathfrak{C}}
\newcommand{\homega}{\widehat{\omega}}
\newcommand{\wcomega}{\mathring{\omega}}
\newcommand{\vzero}{\bvec{0}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\naturals}{\mathbb{N}}
\newcommand{\ip}[3][{}]{\ensuremath{\left \langle #2, #3 \right \rangle_{#1}}}
\newcommand\iid{\stackrel{iid}{\sim}}

\makeatletter
\newcommand*{\ov}[1]{
  \m@th\overline{\mbox{#1}\raisebox{2mm}{}}
}

\def\abs#1{\ensuremath{\left \lvert #1 \right \rvert}}

\let\oldemptyset\emptyset
\let\emptyset\varnothing

\begin{document}

\begin{frontmatter}

\title{Reliable error estimation for Sobol' indices}

\author{Cl\'ementine Prieur, Elise Arnaud, Laurent Gilquin, Fred J. Hickernell, Llu\'{i}s Antoni Jim\'{e}nez Rugama}
\address{U. Josef Fourier, Illinois Institute of Technology}
\begin{abstract}
\end{abstract}

\end{frontmatter}

\section{Introduction}

@@

\section{Backgrounds on Sobol' indices}

\subsection{Definition of Sobol' indices}
%We adopt the same notations introduced by Owen in \cite{Owen}. 
Denote by $f$ a numerical model, by $\vx=(x_1,\dots,x_d)$ its vector of inputs and set $\mathcal{D}=\{1,\dots,d\}$. %The uncertainty on $\vx$ is modeled by a random vector that we suppose uniformly distributed on $\cube$.
Let $u$ be a subset of $\mathcal{D}$, $-u$ its complement and $|u|$ its cardinality. $\vx_u$ represents a point in $[0,1)^u$ with components $x_j, j \in u$. Given two points $\vx$ and $\vx'$, the hybrid point $\vw=(\vx_u:{\vx'}_{-u})$ is defined as $w_j=x_j$ if $j \in u$ and $w_j={x'}_j$ if $j \notin u$.

Let $f$ be the symbol representing the numerical model studied. We assume $f \in \mathcal{L}^2[0,1)^d$. Denote by $\mu$ and $\sigma^2$ the mean and variance of $f$.
The uncertainty on $\vx$ is modeled by random variables such that $\vx \iid \mathcal{U}[0,1)^d$. Recall the Hoeffding decomposition \cite{Hoeffding} of $f$:
\begin{equation}
f(\vx)=\mu+\sum \limits_{u \subseteq \mathcal{D}, u \neq \varnothing} f_u(\vx),
\label{anova}
\end{equation}
where:
\[f_u(\vx)= \int_{[0,1)^{|u|}} f(\vx) d{\vx}_{-u} - \sum \limits_{v \subset u} f_v(\vx).\]
Due to orthogonality, taking the variance of each side in equation (\ref{anova}) leads to the variance decomposition of $f$:
\[ \sigma^2 = \sum \limits_{u \subseteq \mathcal{D}, u \neq \varnothing} \sigma_u^2, \ \text{ with } \ \sigma_u^2=\int_{[0,1)^{|u|}} f_u(\vx)^2 d{\vx_u}.\]
From the variance decomposition of $f$, on can express the following two quantities:
\[\underline{\tau}_u^2 = \sum \limits_{v \subseteq u} \sigma_v^2, \qquad
\ov{$\tau$}_u^2 = \sum \limits_{v \cap u \neq \varnothing} \sigma_v^2, \qquad u \subsetneq \mathcal{D}.\]

@@

The problem of interest here is the evaluation of first-order and total effect Sobol' indices. The computation of these indices is performed based on the following integral formulas for their numerators:
\begin{align}
\label{first.order}
\underline{\tau}_u^2  &= \int_{[0,1)^{2d-1}} \left(f(\vx)-
f(\vx_u:{\vx'}_{-u})\right)f(\vx')d\vx d{\vx'_{-u}}, \\
\label{total.effect}
\ov{$\tau$}_u^2 &= \frac{1}{2}\int_{[0,1)^{d+1}}(f(\vx')-f(\vx_u:{\vx'}_{-u}))^2 d{\vx'} d\vx_u, \qquad u \in \mathcal{D},
\end{align}
The variance of $f$ is evaluated by:
\begin{equation}
\sigma^2 = \int_{[0,1)^{d}} f(\vx)^2d{\vx} - \mu^2, \text{ with } \ \mu = \int_{[0,1)^{d}} f(\vx) d{\vx}.
\label{eq.sigma.mu}
\end{equation} 
Most of the time, the complexity of $f$ causes the computation of integrals (\ref{first.order}), (\ref{total.effect}) and (\ref{eq.sigma.mu}) to be intractable. In such case, one needs to rely on an estimation of these quantities.

\subsection{Estimation of Sobol' indices}
\label{estimation.strategies}
In this section we review two Monte Carlo procedures for the estimation of Sobol' indices. We define by design a point set $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ obtained by sampling each variable $x_j$ $n$ times. Each row of the design is a point $\vx_i$ in $[0,1)^d$. Each column of the design refers to a variable $x_j$. Consider $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P'}=\{{\vx'}_i\}_{i=0}^{n-1}$ two designs where $(\vx_i,{\vx'}_i) \iid [0,1)^{2d}$. 
\bigskip

One way to estimate the two quantities (\ref{first.order}) and (\ref{total.effect}) is via:
\begin{align}
\label{first.order.est}
\widehat{\underline{\tau}_u^2} & = \frac{1}{n} \sum \limits_{i=0}^{n-1} \left(f(\vx_i)-f(\vx_{i,u}:{\vx'}_{i,-u}\right)f(\vx'_i),\\
\label{total.effect.est}
\widehat{\ov{$\tau$}_u^2} & = \frac{1}{2n} \sum \limits_{i=0}^{n-1} (f({\vx'}_i) - f(\vx_{i,u}:{\vx'}_{i,-u}))^2, \qquad u \in \mathcal{D},
\end{align}
using for $\sigma^2$ the classic estimator:
\begin{equation}
 \widehat{\sigma}^2 = \frac{1}{n} \sum \limits_{i=0}^{n-1} f(\vx_i)^2 - \widehat{\mu}^2, \text{ with } \ \widehat{\mu} =  \frac{1}{n} \sum \limits_{i=0}^{n-1} f(\vx_i).
\label{mu.est}
\end{equation}
The Sobol' indices estimators are defined by:
\begin{equation*}
\widehat{\underline{S}_u} = \widehat{\underline{\tau}_u^2} / \widehat{\sigma}^2, \qquad
\widehat{\overline{S}_u} = \widehat{\overline{\tau}_u^2} / \widehat{\sigma}^2.
\end{equation*}
The estimation of a single pair ($\widehat{\underline{S}_u}$, $\widehat{\overline{S}_u}$) requires $3n$ evaluations of the model $f$. Using a combinatorial formalism, in \cite[Theorem 1]{Saltelli} Saltelli proposes the following estimation strategy:
\begin{theorem}
\label{saltelli.theorem}
The $d+2$ designs $\{\vx_{i,u},{\vx'}_{i,-u}\}_{i=0}^{n-1}$ constructed for $u \in \{\varnothing,\{1\},\dots,$ $\{d\},\mathcal{D}\}$ allows to estimate all first-order and all total effect Sobol' indices at a cost of $n(d+2)$ evaluations of the model.
\end{theorem}
The $d+2$ designs of Theorem \ref{saltelli.theorem} are obtained by substituting columns of $\mathcal{P}$ for columns of $\mathcal{P}'$ accordingly to $u$. While elegant, this approach still requires a number of model evaluations that grows linearly with respect to the input space dimension.

An efficient alternative to evaluate all first-order indices was proposed by Mara \textit{et al.} \cite{Mara}  requiring only $2n$ model evaluations. This alternative relies on the construction of two replicated designs. The notion of replicated designs was first introduced by McKay through its introduction of replicated Latin Hypercubes in \cite{Mckay}. The definition we give here introduce the structure in a wider context:
\begin{definition}
\label{rep.designs}
Let $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P}'=\{{\vx'}_i\}_{i=0}^{n-1}$ be two point sets in
$[0,1)^{d}$. Let $\mathcal{P}^u=\{\vx_{i,u}\}_{i=0}^{n-1}$ (resp. ${\mathcal{P}'}^u$), $u \subsetneq \mathcal{D}$, denote the subset of dimensions of $\mathcal{P}$ (resp. $\mathcal{P}'$) indexed by $u$. We say that $\mathcal{P}$ and $\mathcal{P}'$ are two replicated designs of order $a \in \{1,\dots,d-1\}$ if $\forall \ u \subsetneq \mathcal{D}$ such that $|u|=a$, $\mathcal{P}^u$ and ${\mathcal{P}'}^u$ are the same point set in $[0,1)^a$. We note $\pi_u$ the permutation arranging the rows of ${\mathcal{P}'}^u$ into $\mathcal{P}^u$.
\end{definition}
The method introduced in \cite{Mara} allows to estimate all first-order Sobol' indices with only two replicated designs of order $1$. The key point of this method is to use the permutations resulting from the structure of the two replicated designs to mimic the hybrid points in formula (\ref{first.order.est}).

More precisely, let $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P}'=\{{\vx'}_i\}_{i=0}^{n-1}$ be two replicated designs of order $1$. Denote by $\{f(\vx_i)\}_{i=0}^{n-1}$ and $\{f({\vx'}_i)\}_{i=0}^{n-1}$ the two sets of model evaluations obtained with $\mathcal{P}$ and $\mathcal{P}'$. Consider $u \in \mathcal{D}$, from Definition \ref{rep.designs} there exists a permutation $\pi_u$ such that ${\vx'}_{\pi_u(i),u}={\vx}_{i,u}$. Then, remark that $\forall i \in \{0,\dots,n-1\}$:
\[\pi_u(f({\vx'}_i))=f({\vx'}_{\pi_u(i)})=f(\vx'_{\pi_u(i),u}:{\vx'}_{\pi_u(i),-u})=f(\vx_{i,u}:{\vx'}_{\pi_u(i),-u}).\]
Hence, each $\underline{\tau}^2_u$ can be estimated via formula (\ref{first.order.est}) with $\{f(\vx_i)\}_{i=0}^{n-1}$ and $\{\pi_u(f({\vx'}_i))\}_{i=0}^{n-1}$ without requiring further model evaluations. This estimation method has been deeply studied and generalized in Tissot et al. \cite{Mara} to the case of closed second-order indices. In the following we refer to this method as replication procedure.

\subsection{Towards a reliable estimation}
The aim of this paper is to propose a recursive procedure to estimate first-order and total effect Sobol' indices. The choice of the stopping criterion is the key point of this paper. In \cite{Gilquin16}, the authors propose a recursive version of the replication procedure. In this recursive version, the stopping criterion is an absolute difference between estimates computed on consecutive steps.

More generally, in most recursive approaches that estimate Sobol' indices, the stopping criterion is a quantity of interest build directly upon the estimates. Such stopping criteria often involve hyper-parameters hard to tweak but above all fail to guarantee any error bound on the estimates. 
\bigskip

Our recursive procedure stands apart from others approaches with the construction of a robust stopping criterion. This criterion is an error bound based on the discrete Walsh decomposition of integrals (\ref{first.order}) and (\ref{total.effect}). This decomposition use the close set property of digital sequences. As such, our recursive procedure relies on an iterative construction of Sobol' sequences. This construction is performed accordingly to the multiplicative  approach presented in \cite{crass}.

The construction of the error bound is the subject of the next section. Our recursive estimation procedure is presented in section \ref{section.recursive.procedure}.

\section{Reliable error bound for Sobol' indices}
\label{section.error}

\subsection{Reliable integral estimation using quasi-Monte Carlo}
\label{section.error.algo}

We assume we have an embedded sequence of digital nets in base $b$ as in \cite[Sec. 2-3]{HicJim},
\[
\mathcal{P}_0=\{\bvec{0}\}\subset\dots\subset\mathcal{P}_m=\{\vx_i\}_{i=0}^{b^m-1}\subset\dots\subset\mathcal{P}_\infty=\{\vx_i\}_{i=0}^{\infty}.
\]
Each $\mathcal{P}_m$ has a group structure under the digit wise addition:
\[
\vx \oplus \vt = \left(\sum_{\ell=1}^{\infty} [(x_{j\ell} + t_{j\ell}) \bmod b] b^{-\ell} \pmod{1} \right)_{j=1}^d,
\]
where $x_{j\ell}$ and $t_{j\ell}$ are the $b$-adic decompositions of the $j^{\rm th}$ component of points $\vx$ and $\vt$.

In order to relate the group structure of $\mathcal{P}_m$ with the integration error, we introduce the \emph{dual net} which establishes the relationship between any digital net and the \emph{wavenumber} space $\mathbb{N}_0^d$. Hence, a dual net is
\[
\mathcal{P}_m^\perp=\{\vk\in\mathbb{N}_0^d:\ip{\vx}{\vk}=0,\, \vx\in\mathcal{P}_m\},
\]
and inherits the same embedded structure as for the digital nets,
\begin{equation}\label{dual_net_structure}
\mathcal{P}_0^\perp=\left\{\mathbb{N}_0^d\right\}\supset\dots\supset\mathcal{P}_\infty=\{\bvec{0}\}.
\end{equation}

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
%gsave
%newpath
%  20 20 moveto
%  20 220 lineto
%  220 220 lineto
%  220 20 lineto
%closepath
%2 setlinewidth
%gsave
%  .4 setgray fill
%grestore
%stroke
%grestore
%\end{filecontents*}
%
\RequirePackage{fix-cm}
\RequirePackage{amsmath}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
%
\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
\let\amalg\relax
\usepackage{amssymb,
mathtools,bm,extraipa,mathabx,graphicx,algorithm}
\usepackage{algpseudocode}
\usepackage{url}

%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}

\newcommand{\fudge}{\fC}
\newcommand{\cc}{\mathcal{C}}
\newcommand{\dtf}{\textit{\doubletilde{f}}}
\newcommand{\cube}{[0,1)^d}
\newcommand{\rf}{\mathring{f}}
\newcommand{\rnu}{\mathring{\nu}}
\newcommand{\natm}{\naturals_{0,m}}
\newcommand{\wcS}{\widecheck{S}}
\newcommand{\tol}{\text{tol}}
\newcommand{\e}{\text{e}}
%\newcommand{\bvec}[1]{\boldsymbol{#1}}
\newcommand{\vect}[1]{\boldsymbol{\mathbf{#1}}}
\newcommand{\vx}{\vect{x}}
\newcommand{\vI}{\vect{I}}
\newcommand{\vk}{\vect{k}}
\newcommand{\vw}{\vect{w}}
\newcommand{\vz}{\vect{z}}
\newcommand{\vt}{\vect{t}}
\newcommand{\dif}{\mathsf{d}}
\newcommand{\hf}{\hat{f}}
\newcommand{\hS}{\widehat{S}}
\newcommand{\tS}{\widetilde{S}}
\newcommand{\tf}{\tilde{f}}
\newcommand{\fC}{\mathfrak{C}}
\newcommand{\homega}{\widehat{\omega}}
\newcommand{\wcomega}{\mathring{\omega}}
\newcommand{\vzero}{\vect{0}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\naturals}{\mathbb{N}}
\newcommand{\ip}[3][{}]{\ensuremath{\left \langle #2, #3 \right \rangle_{#1}}}
\newcommand\iid{\stackrel{iid}{\sim}}

\makeatletter
\newcommand{\ov}[1]{
  \m@th\overline{\mbox{#1}\raisebox{2mm}{}}
}
\newcommand{\absolute}[1]{\left\lvert#1\right\rvert}
\let\oldemptyset\emptyset
\let\emptyset\varnothing

% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{Reliable error estimation for Sobol' indices%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

\author{Laurent Gilquin \and
        Llu\'{i}s Antoni Jim\'{e}nez Rugama %etc.
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{Laurent Gilquin \at
              Inria Grenoble - Rh\^{o}ne-Alpes, Inovall\'{e}e, 655 avenue de l'Europe, 38330 Montbonnot \\
              %Tel.: +123-45-678910\\
              %Fax: +123-45-678910\\
              \email{gilquin.laurent@inria.fr}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           Llu\'{i}s Antoni Jim\'{e}nez Rugama \at
              Illinois Institute of Technology, Rettaliata Engineering Center, 10 W 32st, Chicago, IL 60616 \\
\email{ljimene1@hawk.iit.edu}
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
Insert your abstract here. Include keywords, PACS and mathematical
subject classification numbers as needed.
\keywords{First keyword \and Second keyword \and More}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

%\section{Introduction}
%\label{intro}
%Your text comes here. Separate text sections with
%\section{Section title}
%\label{sec:1}
%Text with citations \cite{RefB} and \cite{RefJ}.
%\subsection{Subsection title}
%\label{sec:2}
%as required. Don't forget to give each section
%and subsection a unique label (see Sect.~\ref{sec:1}).
%\paragraph{Paragraph headings} Use paragraph headings as needed.
%\begin{equation}
%a^2+b^2=c^2
%\end{equation}

% For one-column wide figures use
%\begin{figure}
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
%  \includegraphics{example.eps}
% figure caption is below the figure
%\caption{Please write your figure caption here}
%\label{fig:1}       % Give a unique label
%\end{figure}
%
% For two-column wide figures use
%\begin{figure*}
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
%  \includegraphics[width=0.75\textwidth]{example.eps}
% figure caption is below the figure
%\caption{Please write your figure caption here}
%\label{fig:2}       % Give a unique label
%\end{figure*}
%
% For tables use
%\begin{table}
%% table caption is above the table
%\caption{Please write your table caption here}
%\label{tab:1}       % Give a unique label
%% For LaTeX tables use
%\begin{tabular}{lll}
%\hline\noalign{\smallskip}
%first & second & third  \\
%\noalign{\smallskip}\hline\noalign{\smallskip}
%number & number & number \\
%number & number & number \\
%\noalign{\smallskip}\hline
%\end{tabular}
%\end{table}

\section{Introduction}
\label{sec:1}

@@

\section{Backgrounds on Sobol' indices}
\label{sec:2}

\subsection{Definition of Sobol' indices}
\label{sec:2.1}
%We adopt the same notations introduced by Owen in \cite{Owen}. 
Denote by $f$ a numerical model, by $\vx=(x_1,\dots,x_d)$ its vector of inputs and set $\mathcal{D}=\{1,\dots,d\}$. %The uncertainty on $\vx$ is modeled by a random vector that we suppose uniformly distributed on $\cube$.
Let $u$ be a subset of $\mathcal{D}$, $-u$ its complement and $|u|$ its cardinality. $\vx_u$ represents a point in $[0,1)^u$ with components $x_j, j \in u$. Given two points $\vx$ and $\vx'$, the hybrid point $\vw=(\vx_u:{\vx'}_{-u})$ is defined as $w_j=x_j$ if $j \in u$ and $w_j={x'}_j$ if $j \notin u$.

Let $f$ be the symbol representing the numerical model studied. We assume $f \in \mathcal{L}^2[0,1)^d$. Denote by $\mu$ and $\sigma^2$ the mean and variance of $f$.
The uncertainty on $\vx$ is modeled by random variables such that $\vx \iid \mathcal{U}[0,1)^d$. Recall the Hoeffding decomposition \cite{Hoeffding,Sobol'} of $f$:
\begin{equation}
f(\vx)=\mu+\sum \limits_{u \subseteq \mathcal{D}, u \neq \varnothing} f_u(\vx),
\label{anova}
\end{equation}
where:
\[f_u(\vx)= \int_{[0,1)^{|u|}} f(\vx) d{\vx}_{-u} - \sum \limits_{v \subset u} f_v(\vx).\]
Due to orthogonality, taking the variance of each side in equation (\ref{anova}) leads to the variance decomposition of $f$:
\[ \sigma^2 = \sum \limits_{u \subseteq \mathcal{D}, u \neq \varnothing} \sigma_u^2, \ \text{ with } \ \sigma_u^2=\int_{[0,1)^{|u|}} f_u(\vx)^2 d{\vx_u}.\]
From the variance decomposition of $f$, on can express the following two quantities:
\[\underline{\tau}_u^2 = \sum \limits_{v \subseteq u} \sigma_v^2, \qquad
\ov{$\tau$}_u^2 = \sum \limits_{v \cap u \neq \varnothing} \sigma_v^2, \qquad u \subsetneq \mathcal{D}.\]

@@

The problem of interest here is the evaluation of first-order and total effect Sobol' indices. The computation of these indices is performed based on the following integral formulas for their numerators:
\begin{align}
\label{first.order}
\underline{\tau}_u^2  &= \int_{[0,1)^{2d-1}} \left(f(\vx)-
f(\vx_u:{\vx'}_{-u})\right)f(\vx')d\vx d{\vx'_{-u}}, \\
\label{total.effect}
\ov{$\tau$}_u^2 &= \frac{1}{2}\int_{[0,1)^{d+1}}(f(\vx')-f(\vx_u:{\vx'}_{-u}))^2 d\vx_u d{\vx'}, \ u \in \mathcal{D},
\end{align}
The variance of $f$ is evaluated by:
\begin{equation}
\sigma^2 = \int_{[0,1)^{d}} f(\vx)^2d{\vx} - \mu^2, \text{ with } \ \mu = \int_{[0,1)^{d}} f(\vx) d{\vx}.
\label{eq.sigma.mu}
\end{equation} 
Most of the time, the complexity of $f$ causes the computation of integrals (\ref{first.order}), (\ref{total.effect}) and (\ref{eq.sigma.mu}) to be intractable. In such case, one needs to rely on an estimation of these quantities.

\subsection{Estimation of Sobol' indices}
\label{sec:2.2}
In this section we review two Monte Carlo procedures for the estimation of Sobol' indices. We define by design a point set $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ obtained by sampling each variable $x_j$ $n$ times. Each row of the design is a point $\vx_i$ in $[0,1)^d$. Each column of the design refers to a variable $x_j$. Consider $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P'}=\{{\vx'}_i\}_{i=0}^{n-1}$ two designs where $(\vx_i,{\vx'}_i) \iid [0,1)^{2d}$. One way to estimate the two quantities (\ref{first.order}) and (\ref{total.effect}) is via:
\begin{align}
\label{first.order.est}
\widehat{\underline{\tau}_u^2} & = \frac{1}{n} \sum \limits_{i=0}^{n-1} \left(f(\vx_i)-f(\vx_{i,u}:{\vx'}_{i,-u})\right)f(\vx'_i),\\
\label{total.effect.est}
\widehat{\ov{$\tau$}_u^2} & = \frac{1}{2n} \sum \limits_{i=0}^{n-1} (f({\vx'}_i) - f(\vx_{i,u}:{\vx'}_{i,-u}))^2, \ u \in \mathcal{D},
\end{align}
using for $\sigma^2$ the classic estimator:
\begin{equation}
 \widehat{\sigma}^2 = \frac{1}{n} \sum \limits_{i=0}^{n-1} f(\vx_i)^2 - \widehat{\mu}^2, \text{ with } \ \widehat{\mu} =  \frac{1}{n} \sum \limits_{i=0}^{n-1} f(\vx_i).
\label{mu.est}
\end{equation}
The Sobol' indices estimators are defined by:
\begin{equation}
\widehat{\underline{S}_u} = \widehat{\underline{\tau}_u^2} / \widehat{\sigma}^2, \qquad
\widehat{\overline{S}_u} = \widehat{\overline{\tau}_u^2} / \widehat{\sigma}^2.
\label{common.sobol.est}
\end{equation}
The estimation of a single pair ($\widehat{\underline{S}_u}$, $\widehat{\overline{S}_u}$) requires $3n$ evaluations of the model $f$. Using a combinatorial formalism, in \cite{Saltelli} Saltelli proposes the following estimation strategy:
\begin{theorem}
\label{saltelli.theorem}
The $d+2$ designs $\{\vx_{i,u},{\vx'}_{i,-u}\}_{i=0}^{n-1}$ constructed for $u \in \{\varnothing,\{1\},\dots,$ $\{d\},\mathcal{D}\}$ allows to estimate all first-order and all total effect Sobol' indices at a cost of $n(d+2)$ evaluations of the model.
\end{theorem}
The $d+2$ designs of Theorem \ref{saltelli.theorem} are obtained by substituting columns of $\mathcal{P}$ for columns of $\mathcal{P}'$ accordingly to $u$. While elegant, this approach still requires a number of model evaluations that grows linearly with respect to the input space dimension.

An efficient alternative to evaluate all first-order indices was proposed by \cite{Mara} requiring only $2n$ model evaluations. This alternative relies on the construction of two replicated designs. The notion of replicated designs was first introduced by McKay through its introduction of replicated Latin Hypercubes \cite{McKay}. The definition we give here introduce the structure in a wider context:
\begin{definition}
\label{rep.designs}
Let $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P}'=\{{\vx'}_i\}_{i=0}^{n-1}$ be two point sets in
$[0,1)^{d}$. Let $\mathcal{P}^u=\{\vx_{i,u}\}_{i=0}^{n-1}$ (resp. ${\mathcal{P}'}^u$), $u \subsetneq \mathcal{D}$, denote the subset of dimensions of $\mathcal{P}$ (resp. $\mathcal{P}'$) indexed by $u$. We say that $\mathcal{P}$ and $\mathcal{P}'$ are two replicated designs of order $a \in \{1,\dots,d-1\}$ if $\forall \ u \subsetneq \mathcal{D}$ such that $|u|=a$, $\mathcal{P}^u$ and ${\mathcal{P}'}^u$ are the same point set in $[0,1)^a$. We note $\pi_u$ the permutation arranging the rows of ${\mathcal{P}'}^u$ into $\mathcal{P}^u$.
\end{definition}
The method introduced in \cite{Mara} allows to estimate all first-order Sobol' indices with only two replicated designs of order $1$. The key point of this method is to use the permutations resulting from the structure of the two replicated designs to mimic the hybrid points in formula (\ref{first.order.est}).

More precisely, let $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P}'=\{{\vx'}_i\}_{i=0}^{n-1}$ be two replicated designs of order $1$. Denote by $\{f(\vx_i)\}_{i=0}^{n-1}$ and $\{f({\vx'}_i)\}_{i=0}^{n-1}$ the two sets of model evaluations obtained with $\mathcal{P}$ and $\mathcal{P}'$. Consider $u \in \mathcal{D}$, from Definition \ref{rep.designs} there exists a permutation $\pi_u$ such that ${\vx'}_{\pi_u(i),u}={\vx}_{i,u}$. Then, remark that $\forall i \in \{0,\dots,n-1\}$:
\begin{align*}
\pi_u(f({\vx'}_i))&=f({\vx'}_{\pi_u(i)}), \\
\pi_u(f({\vx'}_i))&=f(\vx'_{\pi_u(i),u}:{\vx'}_{\pi_u(i),-u}), \\
\pi_u(f({\vx'}_i))&=f(\vx_{i,u}:{\vx'}_{\pi_u(i),-u}).
\end{align*}
Hence, each $\underline{\tau}^2_u$ can be estimated via formula (\ref{first.order.est}) with \\ $\pi_u(f({\vx'}_i))$ in place of $f(\vx_{i,u}:{\vx'}_{i,-u})$ without requiring further model evaluations. This estimation method has been deeply studied and generalized in Tissot et al. \cite{Tissot} to the case of closed second-order indices. In the following we refer to this method as replication procedure.

\subsection{Towards a reliable estimation}
\label{sec:2.3}
The aim of this paper is to propose a recursive procedure to estimate first-order and total effect Sobol' indices. The choice of the stopping criterion is the key point of this paper. %A recursive version of the replication procedure has recently been proposed by Gilquin et al. (2016) (preprint available at \url{https://hal.inria.fr/hal-01291769}). In this recursive version, the stopping criterion is an absolute difference between estimates computed on consecutive steps. 
Generally, in most recursive approaches that estimate Sobol' indices, the stopping criterion is a quantity of interest build directly upon the estimates. Such stopping criteria often involve hyper-parameters hard to tweak but above all fail to guarantee any error bound on the estimates. 
\bigskip

Our recursive procedure stands apart from others approaches with the construction of a robust stopping criterion. This criterion is an error bound based on the discrete Walsh decomposition of integrals (\ref{first.order}) and (\ref{total.effect}). This decomposition use the close set property of digital sequences. As such, our recursive procedure relies on an iterative construction of Sobol' sequences. This construction is performed accordingly to the multiplicative approach presented in (Gilquin et al., 2016, preprint available at \url{https://hal.inria.fr/hal-01349444}).

The construction of the error bound is the subject of the next section. Our recursive estimation procedure is detailed in section \ref{sec:4}.

\section{Reliable error bound for Sobol' indices}
\label{sec:3} 
We start by reviewing the construction of an error bound for the estimation of a multidimensional integral proposed in \cite{HicJim}. This construction is based on the Walsh decomposition of the integrand and relies on quasi-Monte Carlo points sets. Then, we present an extension of this error bound for the case of Sobol' indices. This extension exploits the integral formula of a Sobol' index.

\subsection{Reliable integral estimation using quasi-Monte Carlo}
\label{sec:3.1}

We assume we have an embedded sequence of digital nets in base $b$ as in \cite[Sec. 2-3]{HicJim},
\[
\mathcal{P}_0=\{\vect{0}\}\subset\dots\subset\mathcal{P}_m=\{\vx_i\}_{i=0}^{b^m-1}\subset\dots\subset\mathcal{P}_\infty=\{\vx_i\}_{i=0}^{\infty}.
\]
Each $\mathcal{P}_m$ has a group structure under the digit wise addition:
\[
\vx \oplus \vt = \left(\sum_{\ell=1}^{\infty} [(x_{j\ell} + t_{j\ell}) \bmod b] b^{-\ell} \pmod{1} \right)_{j=1}^d,
\]
where $x_{j\ell}$ and $t_{j\ell}$ are the $b$-adic decompositions of the $j^{\rm th}$ component of points $\vx$ and $\vt$.
